---
title: "♛ rchess A Chess package for R"
output:
  html_document:
    theme: journal
    toc: yes
---
<!-- README.md is generated from README.Rmd -->

```{r setup, echo = FALSE, warning=FALSE}
# devtools::install_github("jbkunst/tufterhandout")
library("rchess")
options(digits = 3, knitr.table.format = "markdown")
knitr::opts_chunk$set(collapse = TRUE, warning = FALSE, fig.width = 5.5, fig.height = 5.5, dpi = 72)
```
<script>
jQuery(document).ready(function() {
    var offset = 220;
    var duration = 500;
    jQuery(window).scroll(function() {
        if (jQuery(this).scrollTop() > offset) {
            jQuery('.back-to-top').fadeIn(duration);
        } else {
            jQuery('.back-to-top').fadeOut(duration);
        }
    });
    
    jQuery('.back-to-top').click(function(event) {
        event.preventDefault();
        jQuery('html, body').animate({scrollTop: 0}, duration);
        return false;
    })
});
</script>
<script>
  (function(i,s,o,g,r,a,m){i['GoogleAnalyticsObject']=r;i[r]=i[r]||function(){
  (i[r].q=i[r].q||[]).push(arguments)},i[r].l=1*new Date();a=s.createElement(o),
  m=s.getElementsByTagName(o)[0];a.async=1;a.src=g;m.parentNode.insertBefore(a,m)
  })(window,document,'script','//www.google-analytics.com/analytics.js','ga');

  ga('create', 'UA-17670833-8', 'auto');
  ga('send', 'pageview');

</script>
<link href="https://fonts.googleapis.com/css?family=Playfair+Display" rel="stylesheet" type="text/css">
<style>
  h1, h2, h3, h4, .TOC > li {
    font-family : "Playfair Display", serif;
    color: #b58863;
  } 
  
  a {
    color: #b58863;
  }
  
  .level2 {
    padding-top: 20px;
  }
  
    
  .back-to-top {
    font-family : "Playfair Display", serif;
    position: fixed;
    bottom: 2em;
    right: 0px;
    text-decoration: none;
    color : #b58863;
    background-color : #f0d9b5;
    font-size: 12px;
    padding: 1em;
    display: none;
  }
  
  .html-widget-static-bound {
    margin-left: auto;
    margin-right: auto;
  }
  
  .main-container {
    max-width: 700px;
    margin-left: auto;
    margin-right: auto;
  }
  
  
</style>

<a href="#" class="back-to-top">Back to Top</a>

<a href="https://github.com/jbkunst/rchess" class="github-corner"><svg width="80" height="80" viewBox="0 0 250 250" style="fill:#151513; color:#fff; position: absolute; top: 0; border: 0; right: 0;"><path d="M0,0 L115,115 L130,115 L142,142 L250,250 L250,0 Z"></path><path d="M128.3,109.0 C113.8,99.7 119.0,89.6 119.0,89.6 C122.0,82.7 120.5,78.6 120.5,78.6 C119.2,72.0 123.4,76.3 123.4,76.3 C127.3,80.9 125.5,87.3 125.5,87.3 C122.9,97.6 130.6,101.9 134.4,103.2" fill="currentColor" style="transform-origin: 130px 106px;" class="octo-arm"></path><path d="M115.0,115.0 C114.9,115.1 118.7,116.5 119.8,115.4 L133.7,101.6 C136.9,99.2 139.9,98.4 142.2,98.6 C133.8,88.0 127.5,74.4 143.8,58.0 C148.5,53.4 154.0,51.2 159.7,51.0 C160.3,49.4 163.2,43.6 171.4,40.1 C171.4,40.1 176.1,42.5 178.8,56.2 C183.1,58.6 187.2,61.8 190.9,65.4 C194.5,69.0 197.7,73.2 200.1,77.6 C213.8,80.2 216.3,84.9 216.3,84.9 C212.7,93.1 206.9,96.0 205.4,96.6 C205.1,102.4 203.0,107.8 198.3,112.5 C181.9,128.9 168.3,122.5 157.7,114.1 C157.9,116.9 156.7,120.9 152.7,124.9 L141.0,136.5 C139.8,137.7 141.6,141.9 141.8,141.8 Z" fill="currentColor" class="octo-body"></path></svg></a><style>.github-corner:hover .octo-arm{animation:octocat-wave 560ms ease-in-out}@keyframes octocat-wave{0%,100%{transform:rotate(0)}20%,60%{transform:rotate(-25deg)}40%,80%{transform:rotate(10deg)}}@media (max-width:500px){.github-corner:hover .octo-arm{animation:none}.github-corner .octo-arm{animation:octocat-wave 560ms ease-in-out}}</style>
[![travis-status](https://api.travis-ci.org/jbkunst/rchess.svg)](https://travis-ci.org/jbkunst/rchess)
[![version](http://www.r-pkg.org/badges/version/rchess)](http://www.r-pkg.org/pkg/rchess)
[![downloads](http://cranlogs.r-pkg.org/badges/rchess)](http://www.r-pkg.org/pkg/rchess)

## Introduction

The `rchess` package is a chess move, generation/validation, piece placement/movement, and check/checkmate/stalemate detection.
 

## Installation

Via standart install:

```{r, eval=FALSE}
install.packages("rchess")
```

Or if you want to be risky you can install the latest development version from github with:

```{r, eval=FALSE}
devtools::install_github("jbkunst/rchess")
```

## Basic Usage

```{r}
chss <- Chess$new()
chss
```

Check the legal next moves:

```{r}
chss$moves()

chss$moves(verbose = TRUE)
```

Make a move:

```{r}
chss$move("a3")
```

We can concate some moves (and a captures)

```{r}
chss$move("e5")$move("f4")$move("Qe7")$move("fxe5")
```


```{r}
plot(chss)
```

Or a ggplot2 version (I know, I need to change the [chess pieces symbols in unicode](https://en.wikipedia.org/wiki/Chess_symbols_in_Unicode); maybe use a chess typeface)

```{r}
plot(chss, type = "ggplot")
```


There are function to get information of actual position:
```{r}
chss$turn()

chss$square_color("h1")

chss$get("e5")

chss$history(verbose = TRUE)

chss$history()

chss$undo()

chss$history()

chss$fen()
```

You can edit the header.
```{r}
chss$header("White", "You")
chss$header("WhiteElo", 1800)
chss$header("Black", "Me")
chss$header("Date", Sys.Date())
chss$header("Site", "This R session")
```

And get the pgn. 
```{r}
cat(chss$pgn())
```

Or plot the board in ascii format:
```{r}
chss$ascii()
```

## Load PGN and FEN

### FEN

```{r}
chssfen <- Chess$new()

fen <- "rnbqkbnr/pp1ppppp/8/2p5/4P3/8/PPPP1PPP/RNBQKBNR w KQkq c6 0 2"

chssfen$load(fen)

plot(chssfen)
```

### PGN

```{r}
pgn <- system.file("extdata/pgn/kasparov_vs_topalov.pgn", package = "rchess")
pgn <- readLines(pgn, warn = FALSE)
pgn <- paste(pgn, collapse = "\n")
cat(pgn)

chsspgn <- Chess$new()

chsspgn$load_pgn(pgn)

cat(chsspgn$pgn())

chsspgn$history()

chsspgn$history(verbose = TRUE)

plot(chsspgn)
```

## Validation Functions

### State validation

```{r}
chss2 <- Chess$new("rnb1kbnr/pppp1ppp/8/4p3/5PPq/8/PPPPP2P/RNBQKBNR w KQkq - 1 3")

```

```{r}
plot(chss2)
```


```{r}
chss2$in_check()

chss2$in_checkmate()
```

### Slatemate validation

```{r}
chss3 <- Chess$new("4k3/4P3/4K3/8/8/8/8/8 b - - 0 78")

plot(chss3)

chss3$in_stalemate()
```

### Three fold repetition

```{r}
chss4 <- Chess$new("rnbqkbnr/pppppppp/8/8/8/8/PPPPPPPP/RNBQKBNR w KQkq - 0 1")

chss4$in_threefold_repetition()

chss4$move('Nf3')$move('Nf6')$move('Ng1')$move('Ng8')

chss4$in_threefold_repetition()

chss4$move('Nf3')$move('Nf6')$move('Ng1')$move('Ng8')

chss4$in_threefold_repetition()

chss4$history()
```

### Insufficient material
```{r}
chss5 <- Chess$new("k7/8/n7/8/8/8/8/7K b - - 0 1")

plot(chss5)

chss5$insufficient_material()
```

## Helpers Functions

There some helper function to get more information 

### History Detail

This functions is a detailed version from the `history(verbose = TRUE)`.

```{r}
chsspgn$history_detail()
```

### Plot a boad via ggplot
You can plot a specific fen vía ggplot:
```{r}
ggchessboard(chsspgn$fen())
```

### Auxiliar Data

There function to retrieve some data which is easier to plot:

#### Pieces
```{r}
rchess:::.chesspiecedata()
```

#### Board
```{r}
rchess:::.chessboarddata()
```

## Data

The package have two data sets:

### FIDE World cups

```{r}
data("chesswc")
str(chesswc)
head(chesswc)
```

### Chess Openings

```{r}
data("chessopenings")
str(chessopenings)
head(chessopenings)
```


## More Details

### Under the hood

This package is basically a wrapper of [chessjs](https://github.com/jhlywa/chess.js) by [jhlywa](https://github.com/jhlywa).

The main parts in this package are:

- V8 package and chessjs javascript library.
- R6 package for the OO system.
- htmlwidget package and chessboardjs javascript library.

Thanks to the creators and maintainers of these packages and libraries.

### Session Info

```{r}
print(sessionInfo())
```