---
title: "Complete rchess tour"
description: >
  Play games, load PGN and FEN positions, inspect game state, and render chess
  boards with the HTML widget or ggplot2.
---

```{r setup, include=FALSE}
library(rchess)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 5.5,
  fig.height = 5.5,
  dpi = 96
)
```

This article updates the complete example that originally lived on the
`gh-pages` branch. It follows a game from its initial position through move
history, FEN and PGN loading, game-state validation, and visualization.

## Start a game

Create a game in the standard starting position and inspect its legal moves:

```{r}
game <- Chess$new()
game$moves()
game$moves(verbose = TRUE)
```

Moves use Standard Algebraic Notation and can be chained:

```{r}
game$move("e4")$move("e5")$move("Nf3")$move("Nc6")
game$history()
game$history(verbose = TRUE)
```

Inspect the current position and active side:

```{r}
game$turn()
game$fen()
game$get("e4")
game$square_color("h1")
```

Undo the latest move or print the board in the console:

```{r}
game$undo()
game$ascii()
```

## Render the board

`plot()` returns an interactive chessboard.js widget by default:

```{r widget}
plot(game)
```

Use `type = "ggplot"` for a static ggplot2 board:

```{r ggplot-board}
plot(game, type = "ggplot")
```

You can also render any FEN position directly:

```{r custom-board}
fen <- "rnbqkbnr/pp1ppppp/8/2p5/4P3/8/PPPP1PPP/RNBQKBNR w KQkq c6 0 2"
ggchessboard(fen, perspective = "black")
```

## Load FEN positions

Create a game from FEN or replace the position of an existing game:

```{r}
fen_game <- Chess$new(fen)
fen_game$turn()
fen_game$moves()

fen_game$load("4k3/4P3/4K3/8/8/8/8/8 b - - 0 78")
fen_game$in_stalemate()
```

## Load PGN games

The package includes the famous Kasparov–Topalov game from Wijk aan Zee 1999:

```{r}
pgn_path <- system.file(
  "extdata/pgn/kasparov_vs_topalov.pgn",
  package = "rchess"
)
pgn <- paste(readLines(pgn_path, warn = FALSE), collapse = "\n")

pgn_game <- Chess$new()
pgn_game$load_pgn(pgn)
pgn_game$get_header()
head(pgn_game$history(verbose = TRUE))
```

Detailed history follows each original piece through the game and records
captures:

```{r}
detail <- pgn_game$history_detail()
detail
```

## Add PGN metadata

Headers can be added before exporting a game:

```{r}
annotated <- Chess$new()
annotated$move("e4")$move("e5")
annotated$header("White", "Player One")
annotated$header("Black", "Player Two")
annotated$header("Site", "R session")
annotated$get_header()
cat(annotated$pgn())
```

## Validate game state

Checkmate:

```{r}
mate <- Chess$new(
  "rnb1kbnr/pppp1ppp/8/4p3/5PPq/8/PPPPP2P/RNBQKBNR w KQkq - 1 3"
)
mate$in_check()
mate$in_checkmate()
```

Threefold repetition:

```{r}
repetition <- Chess$new()
repetition$move("Nf3")$move("Nf6")$move("Ng1")$move("Ng8")
repetition$move("Nf3")$move("Nf6")$move("Ng1")$move("Ng8")
repetition$in_threefold_repetition()
```

Insufficient material:

```{r}
material <- Chess$new("k7/8/n7/8/8/8/8/7K b - - 0 1")
material$insufficient_material()
```

## Included data

`rchess` includes opening classifications and FIDE World Cup games:

```{r}
data("chessopenings", package = "rchess")
data("chesswc", package = "rchess")

head(chessopenings)
dplyr::count(chesswc, event)
```

## Under the hood

`rchess` combines the bundled chess.js rules engine with V8 and an R6 API.
Boards are rendered either with the chessboard.js HTML widget or with ggplot2.
This keeps game logic, state inspection, and visualization available through a
single R object.
